Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-22 paper/paper.tex
reshuffle in the introduction

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embedded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and numerical computational
complexity restrictions. Therefore compression methods for achieving high model sparsity
and numerical efficiency without losing much in performance are especially relevant.

Signal processing, specifically, audio or radio frequency signal analysis is inherently
complex-valued, yet $\cplx$-valued models and neural networks, and, in particular, compression
methods tailored to $\cplx$-domain remain a niche field of research. Computational efficiency
reduction methods such as quantization, integer-based arithmetic \citep{citation_needed},
and distillation \citep{distillation} appear to be immediately applicable to $\cplx$-field.
%
Some model sparsification methods, e.g. magnitude pruning \citep{zuo_compression_2019}, which
reactivates zeroed parameters based on gradients, and sparsity inducing regularizers, can be
adapted directly to $\cplx$-valued models. Other methods -- require additional considerations.

\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015}, $\ell_0$
regularization \citep{louizos_learning_2017} and magnitude pruning on large-scale models.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning on the studied architectures, overall pruning is preferable due to
simplicity and speed. They also conclude that, although dropout and $\ell_0$ yield less
consistent performance and are more complex, they induce non-uniform sparsity throughout
the model, which is essential for better compression \verify{(He et al., 2018)}.
%
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%
Therefore it appears that Bayesian and stochastic sparsification techniques can be useful
for models in $\cplx$ domain. \verify{$\ell_0$ -- stoch?}

% on sparsification
Solving the storage and compute constraints encourages development of model compression
and sparsification methods, that focus on favourable tradeoff between performance and size.
%
Most popular techniques include
model quantization from IEEE754,
and dropout.

The key assumption is that the more explicitly zero parameters there is,
the less computations is needed.

\citet{monning_evaluation_2018} state,
that $\cplx$-networks have less intuitive architectures, inherently higher numerical
complexity due to multiplication in $\cplx$, that requires 3-4 floating point multiplications,
dependening on the implementation.

% due to them not being true cplx and just paired real computational graphs

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, thereby breaking up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
has shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated mutlitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}

% out contribution
Inspired by variational sparsification techniques we develop bayesian dropout for $\cplx$-vlaued
networks.

@@

% complex nets and related stuff
data that is naturally complex-valued
studies have shown a potentially richer representation of complex nets
\citep{hirose2003,fijulamin2009}
% [3] Akira Hirose, Complex-valued neural networks: theories and applications, vol. 5, World Scientific, 2003.
% [4] Md Faijul Amin and Kazuyuki Murase, “Single-layered complex-valued neural network for real-valued classification problems,” Neurocomputing, vol. 72, no. 4-6, pp. 945–955, 2009.

@@

Application of complex-valued networks is not new, \citep{citation_needed}. More recently
they were applied to music transcription and speech recognition \citep{trabelsi_deep_2017},
and \citep{yang_complex_2019}. The key motivation behind adoption of $\cplx$-networks is
pervasiveness of $\cplx$-valued data in digital signal processing \citep{citation_needed},
acoustic and radio signal analysis \citep{citation_needed}, and quantum systems modelling
\citep{citation_needed}, etc. All these applications require compliance with algebraic
operations in the field of complex numbers.

@@

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}
\citep{titsias_doubly_2014}

>>>>

% general intro text with motivation
Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. With continual miniaturization ever more
applications can be found in embedded systems. Common embedded applications include
on-device image recognition and signal processing. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge due to storage, real-time throughput, and numerical computational
complexity restrictions. Therefore compression methods for achieving high model sparsity
and numerical efficiency without losing much in performance are especially relevant.
% Solving the storage and compute constraints encourages development of model compression
% and sparsification methods, that focus on favourable tradeoff between performance and size.

Signal processing, specifically, audio or radio frequency signal analysis is inherently
complex-valued, yet $\cplx$-valued models and neural networks, and, in particular, compression
methods tailored to $\cplx$-domain remain a niche field of research. Computational efficiency
reduction methods such as quantization, integer-based arithmetic \citep{citation_needed},
and distillation \citep{distillation} appear to be immediately applicable to $\cplx$-field.
%
Some model sparsification methods, e.g. magnitude pruning \citep{zuo_compression_2019}, which
reactivates zeroed parameters based on gradients, and sparsity inducing regularizers, can be
adapted directly to $\cplx$-valued models. Other methods require additional considerations.
% the more explicitly zero parameters there is, the less computations is needed.

%%% >>>> (cplx)
% complex nets and related stuff
% cplx domain models
% (cplx-transformer) yang-ma-li.tex
% (cplx capacity) monning-manandhar.tex
\citet{monning_evaluation_2018} state,
that $\cplx$-networks have less intuitive architectures, inherently higher numerical
complexity due to multiplication in $\cplx$, that requires 3-4 floating point multiplications,
dependening on the implementation.
% due to them not being true cplx and just paired real computational graphs

@@

data that is naturally complex-valued
studies have shown a potentially richer representation of complex nets
\citep{hirose2003,fijulamin2009}
% [3] Akira Hirose, Complex-valued neural networks: theories and applications, vol. 5, World Scientific, 2003.
% [4] Md Faijul Amin and Kazuyuki Murase, “Single-layered complex-valued neural network for real-valued classification problems,” Neurocomputing, vol. 72, no. 4-6, pp. 945–955, 2009.

@@

Application of complex-valued networks is not new, \citep{citation_needed}. More recently
they were applied to music transcription and speech recognition \citep{trabelsi_deep_2017},
and \citep{yang_complex_2019}. The key motivation behind adoption of $\cplx$-networks is
pervasiveness of $\cplx$-valued data in digital signal processing \citep{citation_needed},
acoustic and radio signal analysis \citep{citation_needed}, and quantum systems modelling
\citep{citation_needed}, etc. All these applications require compliance with algebraic
operations in the field of complex numbers.
%%% <<<< (cplx)

% on dropout
Dropout, \citep{hinton_improving_2012}, prevents overfitting by injecting multiplicative
binary noise into layer's weights, thereby breaking up co-adaptations that could occur
during training. \citet{wang_fast_2013} argue that the overall effect of binary dropout
on the intermediate outputs via the Central Limit Theorem can be approximated by a Gaussian
with weight-input dependent mean and variance. \citet{srivastava_dropout_2014} propose
using independent $\mathcal{N}(1,1)$ noise, arguing that higher entropy of a Gaussian has
better regularizing effect. In multitask learning \citet{cheung_superposition_2019}
has shown the possibility of storing task-specific parameters in non-destructive superposition
within a single network. Their argument implies that if the single task setting is viewed
as replicated mutlitask learning, then by sampling uncorrelated weight masks dropout acts
as superposition method, which better utilizes the learning capacity of the network.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015}, $\ell_0$
regularization \citep{louizos_learning_2017} and magnitude pruning on large-scale models.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning on the studied architectures, overall pruning is preferable due to
simplicity and speed. They also conclude that, although dropout and $\ell_0$ yield less
consistent performance and are more complex, they induce non-uniform sparsity throughout
the model, which is essential for better compression \verify{(He et al., 2018)}.
%
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%
Therefore it appears that Bayesian and stochastic sparsification techniques can be useful
for models in $\cplx$ domain. \verify{$\ell_0$ -- stoch?}

% our contribution
Inspired by variational sparsification we develop bayesian dropout for $\cplx$-valued
networks.

@@

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \citep{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid g(\varepsilon_{l}; \theta))
        % \log p_{\phi}(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased sampling estimators, \citep{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
% (DSVI) titsias-lazaro.tex
\todo{DSVI}
\citep{titsias_doubly_2014}
